---
title: 'REPRODUCIBLE RESEARCH: Assignment 1'
output:
  html_document:
    keep_md: yes
---
###LOADING THE DATA

I downloaded the data and loaded them using R Studio. After that I checked the data format and summarized them. 
```{r}
activity <- read.csv("~/R/Repreducible research/repdata-data-activity/activity.csv")
head(activity)
summary(activity)
```


###MEAN AND MEDIAN TOTAL NUMBER OF STEPS TAKEN PER DAY

Firstly, I created new dataset that contains only complete observations from original one. Then I computed total number steps of each day using `tapply()` function. Finally, I computed mean and median total number of steps per day and plotted the histogram.
```{r}
dat1<-activity[complete.cases(activity),]
step.sum<-tapply(dat1$step,dat1$date,sum)
mean(step.sum,na.rm=TRUE)
median(step.sum,na.rm=TRUE)
hist(step.sum,seq(0,25000,1250),col="orange",xlab="Steps per day",main="")
```


The mean is `r mean(step.sum,na.rm=TRUE) ` and median is `r median(step.sum,na.rm=TRUE)`


###AVERAGE DAILY ACTIVITY PATTERN

I used the modified dataset from previous question. The function `tapply()` was used to compute average steps for each interval. Subsequently, I created time series plot showing average steps by time interval. 
```{r}
int.mean<-tapply(dat1$step,as.factor(dat1$interval),mean)
library(nnet)
int.mean[which.is.max(int.mean)]
plot(int.mean,type="l",ylab="Mean steps ",main="",xlab="Interval")
```


The highest average steps are `r int.mean[which.is.max(int.mean)]` in interval `r names(int.mean[which.is.max(int.mean)])`.


###MISSING VALUES

I decided to replace missing values by average steps of interval. As a first step, I calculated average steps for each interval. Then I converted the output to dataframe and merged it with original dataframe, matching the intervals.. After that, I easilly replaced NA with matched interval average. In the end, I computed mean, median and plotted the histogram. 
 
```{r}
dat2<-activity
my.mean<-function(x) mean(x,na.rm=TRUE)
int.mean2<-tapply(activity$step,as.factor(activity$interval),my.mean)
df.int.mean2<-data.frame(interval=names(int.mean2),mean=as.vector(int.mean2))
dat2<-merge(dat2,df.int.mean2,by="interval")
dat2$steps[is.na(dat2$steps)]<-dat2$mean[is.na(dat2$steps)]
step.sum3<-tapply(dat2$step,dat2$date,sum)
mean(step.sum3,na.rm=TRUE)
median(step.sum3,na.rm=TRUE)
hist(step.sum3,seq(0,25000,1250),col="orange",xlab="Steps per day",main="")
```


In the original dataframe, there is `r nrow(activity)-sum(complete.cases(activity))` NAs.
After replacing NA with interval averages I computed that mean total number of steps is `r mean(step.sum3,na.rm=TRUE)` and median total number of steps is `r median(step.sum3,na.rm=TRUE)`.


###DIFFERENCES IN ACTIVITY PATTERNS BETWEEN WEEKDAYS AND WEEKENDS

I used `strptime()` function to identify days. After that, I splitted the dataset to weekend and weekday datasets. Finally, interval means were computed and graph plotted for each dataset.

```{r}
Rdate2<-strptime(as.character(dat2$date),"%Y-%m-%d")
dat3<-cbind(dat2,wd=as.factor(Rdate2$wday))
levels(dat3$wd)<-c("weekend","weekday","weekday","weekday","weekday","weekday","weekend")
dat.weekend<-dat3[dat3$wd=="weekend",]
dat.weekday<-dat3[dat3$wd=="weekday",]
int.mean.weekend<-tapply(dat.weekend$step,as.factor(dat.weekend$interval),mean)
int.mean.weekday<-tapply(dat.weekday$step,as.factor(dat.weekday$interval),mean)
plot(int.mean.weekend,type="l",ylab="Mean steps ",main="Weekends",xlab="Interval")
plot(int.mean.weekday,type="l",ylab="Mean steps ",main="Weekdays",xlab="Interval")

```


